---
title: "Homework 11"
author: "Sania Mir"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above. For any included plots, make sure they are clearly labeled. You are free to use any plot type that you feel best communicates the results of your analysis.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
``` 

```{r}
options(scipen = 999)
```

## Data
For this homework, we are going to use the [Gapminder](https://www.gapminder.org/data/) data on population (pop), GDP per capita (gdp), and life expectancy (lex). These data were downloaded and compiled using the `gapminder_compile.Rmd` script in the data folder.  

Let's focus our attention on data from the past 100 years, between 1925 and 2025.
```{r message=FALSE, warning=FALSE}
gapminder <- read_csv("data/gapminder.csv") %>% 
  clean_names() %>% 
  filter(between(year, 1925, 2025))
```

1. Use the function(s) of your choice to get an idea of the overall structure of the data, including its dimensions, column names, classes, etc. 
```{r}
glimpse(gapminder)

summary(gapminder)
```

2. How many countries are represented in this dataset?
```{r}
gapminder %>% 
  count(country)
```

3. Which five countries have had the largest absolute population growth between 1925-2025? Show this as a table.
```{r}
pop_growth <- gapminder %>% 
  filter(year %in% c(1925, 2025)) %>% 
  select(country, year, pop) %>% 
  pivot_wider(names_from = year,values_from = pop,names_prefix = "pop_") %>% 
  mutate(pop_growth = pop_2025 - pop_1925) %>% 
  slice_max(pop_growth, n = 5)

pop_growth
```

4. Make a plot that shows population growth over time for the top country you found in question #3.
```{r}
gapminder %>% 
  filter(country == "India") %>%
  ggplot(aes(x = year, y = pop)) +
  geom_line(color = "blue") +
  labs(title = "population growth over time",x = "yr",y = "pop") +
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 60, hjust = 1))
```

5. Has global life expectancy changed between 1925 and 2025? Show the min, median, mean, and max for all countries (combined) in the dataset.
```{r}
gapminder %>% 
  summarize(min_lex = min(lex, na.rm = T),median_lex = median(lex, na.rm = T),mean_lex = mean(lex, na.rm = T),max_lex = max(lex, na.rm = T))
#Yes global life expectancy has changed significantly between 1925 and 2025. shows that the minimum life expectancy increased and the median and mean life expectancy are much higher in 2025, indicating that most countries now have life expectancy around 60 years or higher, max life expectancy is high
```

6. Make a plot that shows the distribution of life expectancy for all countries in 1925 and 2025 (try using `geom_density`). What do you notice about the shift in life expectancy over time?
```{r}
gapminder %>% 
  filter(year==1925|year==2025) %>% 
  ggplot(aes(x=lex,fill=as.factor(year))) +
  geom_density(alpha=0.5) +
  labs(title="life expectancy for all countries in 1925 and 2025",x = "life expectancy",fill = "yr") +
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 60, hjust = 1))
#shows that life expectancy has shifted to the right from 1925 to 2025 indicates that overall countries have higher life expectancy in 2025 compared to 1925 bc significant improvements in global health
```

7. We are interested in the relationship between per capita GDP and life expectancy; i.e. does having more money help you live longer in 2025? Show this as a plot. (Suggestion! Remove Monaco as an extreme outlier.)
```{r}
gapminder %>% 
  filter(year==2025,country!= "Monaco") %>% 
  ggplot(aes(x=gdp,y=lex))+
  geom_point()+
  labs(title = "relationship between per capita GDP and life expectancy in 2025",x="per capita GDP",y="life expectancy")+
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 60, hjust = 1))
#shown that money indeed does help you live longer, shown by the linear increase 
```

8. Which five countries have had the highest absolute GDP per capita growth over the past 50 years? Show this as a table. (Suggestion! Remove Monaco as an extreme outlier.)
```{r}
gdp_growth<-gapminder %>% 
  filter(year%in%c(1975,2025),country!="Monaco")%>% 
  select(country,year,gdp)%>% 
  pivot_wider(names_from=year,values_from=gdp,names_prefix="gdp_")%>% 
  mutate(gdp_growth=gdp_2025-gdp_1975)%>% 
  slice_max(gdp_growth,n=5)
gdp_growth
```

9. How does per capita GDP growth compare between these same five countries over the past 50 years? Show this as a plot.
```{r}
gapminder %>% 
  filter(country%in% gdp_growth$country,year>= 1975) %>% 
  ggplot(aes(x=year,y=gdp,color=country)) +
  geom_line() +
  labs(title = "per capita GDP growth in same 5 countries over the past 50 years",x = "yr",y = "GDP per Capita",color="Country") +
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 60, hjust = 1))
#all countries show an overall increase over the part 50 yrs, Luxembourg and Norway is stabilizing very slowly increasing
```

10. Do one analysis of your choice that includes a table and plot as outputs.

Table: Average GDP per year
```{r}
gapminder
```

```{r}
avg_gdp <- gapminder %>% 
  group_by(year) %>% 
  summarize(mean_gdp=mean(gdp,na.rm=T))
avg_gdp
```

```{r}
avg_gdp %>% 
  ggplot(aes(x=year,y=mean_gdp))+
  geom_line()+
  labs(title="ave Global GDP per Capita",x = "yr",y = "ave GDP per Capita") +
  theme(legend.position = "bottom",axis.text.x = element_text(angle = 60, hjust = 1))

#The average global GDP per capita has increased  between 1925 and 2025 indicating overall economic growth
```

## Knit and Upload
Please knit your work as an .html file and upload to Canvas. Homework is due before the start of the next lab. No late work is accepted. Make sure to use the formatting conventions of RMarkdown to make your report neat and clean! 